---
title: "Midterm Project Code Draft"
author: "Emma Huizenga"
date: "October ??, 2023"
output: pdf_document
---

You should include all of your code for your midterm project here.

I've included a few chunks of code to help you get started, including:

- loading the tidyverse package, which includes dplyr, ggplot2, and some other useful packages
- an updated ggplot2 theme that approximates the Economist Graphic Detail figures. Feel free to change this or use a different theme!
- code to read in the CCC data
- a final chunk to save your figure

```{r}
library(tidyverse)
```


```{r}
library(ggplot2)
library(RColorBrewer)
library(scales)
library(grid)
library(ggtext)

theme_econ_andy <- function() {
  # Generate the colors for the chart procedurally with RColorBrewer
  palette <- brewer.pal("Greys", n=9)
  color.background = palette[1]
  color.grid.minor = palette[2]
  color.grid.major = palette[3]
  color.axis.text = palette[8] #[6]
  color.axis.title = palette[8]
  color.legend.title = palette[8]
  color.title = palette[9]
  
  grey_8 <- palette[8]
  
  # Begin construction of chart
  theme_bw(base_size=9)  + 
    
    theme(panel.background=element_rect(fill=color.background, color=color.background))  + 
    theme(plot.background=element_rect(fill=color.background, color=color.background))  + 
    theme(panel.border=element_rect(color=color.background))  + 
    
    # Format the grid
    theme(panel.grid.major.y = element_line(color=palette[6], size=.04))  + 
    ## NOTE: Some Economist figures don't include vertical grid lines.
    # To remove them, add + theme(panel.grid.major.x =  element_blank()) to
    # your ggplot plot
    theme(panel.grid.major.x = element_line(color=palette[6], size=.04))  + 
    theme(panel.grid.minor=element_line(color=color.grid.minor, size=.03))  + 
    theme(panel.grid.minor.y=element_blank())  + 
    theme(axis.ticks.y =element_blank())  + 
    theme(axis.ticks.x =element_line(color="black", size=.11))  +
    theme(axis.line.x = element_line(color="black", size=.11))  +
    # remove distance between plotting area and axis:
    theme(panel.border = element_blank()) +
    
    # Format the legend
    theme(legend.background = element_rect(fill=color.background))  + 
    theme(legend.key = element_rect(fill=color.background))  + 
    theme(legend.text = element_text(size=7,color=color.legend.title))  + 
    theme(legend.title = element_text(size=8,color=color.legend.title))  + 
    
    # Set title and axis labels, and format these and tick marks
    theme(plot.title=element_markdown(color=color.title, size=10, vjust=1.25))  + 
    theme(axis.text.x=element_markdown(size=7,color=color.axis.text))  + 
    theme(axis.text.y=element_markdown(size=7,color=color.axis.text))  + 
    theme(axis.title.x=element_markdown(size=8,color=color.axis.title, vjust=0))  + 
    theme(axis.title.y=element_markdown(size=8,color=color.axis.title, vjust=1.25))  + 
    
    # facet stuff
    theme(strip.text = element_text(size=8,color=color.axis.text)) +
    theme(strip.background = element_rect(colour = color.background, 
                                          fill = palette[2], size = 0.5)) +
    # Title fonts
    #theme(plot.title=element_text(family="OpenSans-CondensedBold", margin=margin(b=15)))  + 
    #theme(plot.subtitle=element_text(family="OpenSans-CondensedLightItalic"))  + 
    
    # Plot margins
    theme(plot.margin = unit(c(0.35, 0.2, 0.3, 0.35), "cm"))
}

# Set the theme as the default, so it'll get added to all your figures.
theme_set(theme_econ_andy())
```


```{r}
ccc <- read.csv("ccc_compiled_2023-10-11.csv")
```


Remember that you can combine separate figures into a big figure using patchwork. See https://patchwork.data-imaginist.com/ or the Lecture 4 notes for details.

```{r}
library(patchwork)
```

```{r}
ccc <- ccc %>% 
  select(-starts_with("source"))

mi_data <- ccc %>% 
  filter(state == "MI") %>% 
  mutate(rj = str_detect(issues, "reproductive"))

ccc <- ccc %>% 
  mutate(rj = str_detect(issues, "reproductive"))

ccc$date <- as.Date(ccc$date)
ccc$month <- as.Date(paste(year(ccc$date), month(ccc$date),
                           "01", sep = "-"))

```

```{r}
us_rj <- ccc %>% 
  filter(rj) %>% 
  group_by(month, rj, valence) %>% 
  summarize(count = n())
```


```{r}
us_rj$highlight <- NA
us_rj[us_rj$month >= as.Date("2022-05-02"), "highlight"] <- "yes"

```

```{r}

p1 <- ggplot(us_rj, aes(
    x = month, 
    y = count, 
    fill = factor(highlight)
  )) +  
  geom_col(width = 25) +
  labs(x = "Month", 
       y = "Number of Protests", 
       title = "Protests in the US Over Time", 
       subtitle = "Dobbs Draft Opinion Leak Highlighted in Red") +
  theme_classic() +
  scale_y_continuous(labels = scales::label_comma()) +
  
  theme(legend.position = "none") +
  geom_vline(xintercept = as.Date("2022-06-024"), color = "black", size = .75, alpha = .75, 
             linetype = 2) +
  annotate("text",
         label = "Roe v. Wade \n Overturned",
         x = as.Date("2022-11-11"),
         y = 1400,
         color = "black",
         size = 3, 
         alpha = .9) +
  geom_vline(xintercept = as.Date("2021-12-01"), color = "black", size = .75, alpha = .75, 
             linetype = 2) +
  annotate("text",
         label = "Dobbs v. Jackson \n appealed to Supreme Court",
         x = as.Date("2019-11-03"),
         y = 1300,
         color = "black",
         size = 3, 
         alpha = .9) +
  geom_segment(aes(x = as.Date("2020-08-03"), 
                   xend = as.Date("2021-12-01"), 
                   y = 1300, 
                   yend = 1350, 
                   alpha = .75),
                   color = "black", size = 0.75) +
  geom_vline(xintercept = as.Date("2020-10-26"), color = "black", size = .75, alpha = .75, 
             linetype = 2) +
  annotate("text",
           label = "Amy Coney Barett \n Sworn in",
           x = as.Date("2020-03-01"),
           y = 800,
           color = "black",
           size = 3, 
           alpha = 0.9) +
  geom_vline(xintercept = as.Date("2018-10-6"), color = "black", size = .75, alpha = .75, 
             linetype = 2) +
  annotate("text",
           label = "Brett Kavanaugh \n Sworn in",
           x = as.Date("2018-04-01"),
           y = 600,
           color = "black",
           size = 3, 
           alpha = 0.9) +
  geom_vline(xintercept = as.Date("2017-04-07"), color = "black", size = .75, alpha = .75,
             linetype = 2) +
  annotate("text",
           label = "Neil Gorsuch \n Sworn in",
           x = as.Date("2017-10-01"),
           y = 1100,
           color = "black",
           size = 3, 
           alpha = 0.9) 
```

```{r}

ccc$year <- year(ccc$date)

us_data <- ccc %>% 
  filter(rj) %>% 
  group_by(valence, year) %>% 
  summarize(count = n(), 
            size_mean = mean(size_mean, na.rm = TRUE)) %>% 
  filter(!is.na(valence))

us_data[is.na(us_data$size_mean), "size_mean"] <- 0

valence_lab <- c("Neither", "Left", "Right", NA)
names(valence_lab) <- c("0", "1", "2", NA)

left_data <- subset(us_data, valence == 1)

p2 <- ggplot(left_data, 
       aes(x = year, 
           y = size_mean, 
           fill = factor(valence)
             )) +
  geom_col(fill = "blue",
           color = "black") + 
  labs(
    x = "Year", 
    y = "Average Size of Protests",
    subtitle = "Protests Led by Anti-Trump or Left-Winged Organizations", 
    fill = "Valence") +
  theme_classic() +
  theme(legend.position = "none") 
```


```{r}
mi_data$month <- as.Date(paste(year(mi_data$date), month(mi_data$date),
                           "01", sep = "-"))
mi_data <- mi_data %>% 
     mutate(scaled_size_mean = scale(size_mean))

mi_rj <- mi_data %>% 
  filter(rj) %>% 
  group_by(month, locality) %>%
  summarize(count = n(), 
            size_mean = mean(size_mean, na.rm = TRUE)) 

mi_rj <- mi_rj %>%
  mutate(label = ifelse(!is.na(size_mean) & size_mean > 1000, as.character(locality), NA),
         fill = ifelse(!is.na(size_mean) & size_mean > 1000, "big", "small")) %>%
  mutate(fill = as.factor(fill))



p3 <- ggplot(mi_rj, 
       aes(x = month,
           y = count, 
           alpha = fill,
           label = label, 
           color = label
       )) +
  geom_point(aes(size = size_mean)) +
  labs(x = "Month", 
       y = "Number of Protests", 
       subtitle = "Uprisings in Michigan",
       size = "Average Size") +
  theme_classic() + 
  scale_alpha_manual(values=c(.1, 1)) +
  ggrepel::geom_text_repel() +
  guides(alpha = FALSE,
         color = FALSE) +
  theme(legend.position = "bottom") +
  theme(legend.box.background = 
          element_rect(color = "black", size = 1)) +
  geom_vline(xintercept = as.Date("2022-06-024"), color = "black", alpha = .75) +
  annotate("text",
         label = "Roe v. Wade Overturned",
         x = as.Date("2022-05-09"),
         y = 8,
         color = "black",
         size = 3.5, 
         alpha = .9)
  

```

```{r}
combo_plot <- p1 | (p2 / p3)


combo_plot <- combo_plot + 
  plot_annotation(caption = "Data from the Crowd Counting Consortium",
                  title = "Analyzing the Rise of Reproductive Justice During Contemporary Times")

 combo_plot <- combo_plot + 
   plot_annotation(tag_levels = c('A', '1'), tag_prefix = 'Fig. ', tag_sep = '.', 
                  tag_suffix = ':') &
  theme(plot.tag.position = c(0, 1),
        plot.tag = element_text(size = 10, hjust = 0, vjust = 0))
 
 

```



Finally, you'll want to save your final plot to a pdf so you can load it into the project writeup Rmd document. To do that, you can use the `ggsave` function. You'll need to experiment which changing the size of the image to make it look good in your writeup.

```{r}
ggsave(combo_plot, file="midterm_project_plot.pdf", width=10, height=6)
```

